---
title: "Model-2 DengAI 2017"
author: "Seema Rani Kanuri"
date: "May 14, 2017"
output: html_document
---

#Model : 2

## Introduction : DengAI 2017
Our goal is to predict the total_cases label for each (city, year, weekofyear) in the test set. There are two cities, San Juan and Iquitos, with test data for each city spanning 5 and 3 years respectively. You will make one submission that contains predictions for both cities. The data for each city have been concatenated along with a city column indicating the source: sj for San Juan and iq for Iquitos. The test set is a pure future hold-out, meaning the test data are sequential and non-overlapping with any of the training data. Throughout, missing values have been filled as NaNs.

### Goal:
The Goal is to DengAI predictions for  San Juan and Iquitos city.The goal of the competition is to develop a prediction model that would be able to anticipate the cases of dengue in every country depending on a set of climate variables mentioned above.

```{r}
# Load of libraries

library(tidyverse)
library(zoo)
library(corrplot)
library(MASS)
library(reshape2)
```

### Code with clean tasks (transform NA values, remove of columns.) and exploratory analyses.



```{r}
# Load data

train_features <- read.csv("F:/OneDrive - Texas Tech University/MastersDocuments/DS-Predictive Analytics/DengAI/DataSet/dengue_features_train.csv")

train_labels <- read.csv("F:/OneDrive - Texas Tech University/MastersDocuments/DS-Predictive Analytics/DengAI/DataSet/dengue_labels_train.csv")

test_features <- read.csv("F:/OneDrive - Texas Tech University/MastersDocuments/DS-Predictive Analytics/DengAI/DataSet/dengue_features_test.csv")

submission_format <- read.csv("F:/OneDrive - Texas Tech University/MastersDocuments/DS-Predictive Analytics/DengAI/DataSet/submission_format.csv")
```


```{r}
# Filter of data by city  

sj_train_labels <- filter(train_labels, city == 'sj')
sj_train_features <- filter(train_features, city == 'sj')

iq_train_labels <- filter(train_labels, city == 'iq')
iq_train_features <- filter(train_features, city == 'iq')
```

```{r}
# NA values?

df_na_sj <- as.data.frame(apply(sj_train_features,2, function(x) any(is.na(x))))
colnames(df_na_sj) <- 'is_there_NA'
df_na_sj$number_NA <- apply(sj_train_features,2, function(x) sum(is.na(x)))
df_na_sj$mean_NA <- apply(sj_train_features, 2, function(x) mean(is.na(x)))
```


```{r}
df_na_iq <- as.data.frame(apply(iq_train_features, 2, function(x) any(is.na(x))))
colnames(df_na_iq) <- 'is_there_NA'
df_na_iq$number_NA <- apply(iq_train_features, 2, function(x) sum(is.na(x)))
df_na_iq$mean_NA <- apply(iq_train_features, 2, function(x) mean(is.na(x)))
```

# Vegetation Index over Time Plot with NAs
```{r}
ggplot(sj_train_features, aes(x = as.Date(week_start_date), y = ndvi_ne )) +
  ggtitle('Vegetation Index over Time') +
  theme_bw() +
  xlab('Title') +
  geom_line(na.rm = FALSE, color = 'blue') +
  theme(plot.title = element_text(hjust = 0.5))
```

# Remove 'weekofyear' column

```{r}
sj_train_features <- dplyr::select(sj_train_features, -week_start_date)
iq_train_features <- dplyr::select(iq_train_features, -week_start_date)
```


# Fill of NA values with the previous value
```{r}

sj_train_features <- sj_train_features %>%
            do(na.locf(.))

iq_train_features <- iq_train_features %>%
            do(na.locf(.))
```


# Distribution of labels

```{r}
print(mean(sj_train_labels$total_cases))
print(var(sj_train_labels$total_cases))

print(mean(iq_train_labels$total_cases))
print(var(iq_train_labels$total_cases))
```



```{r}
ggplot(sj_train_labels, aes(x = total_cases)) +
  theme_bw() +
  ggtitle('Cases of dengue in San Juan') +
  geom_histogram() +
  theme(plot.title = element_text(hjust = 0.5))
```


```{r}
ggplot(iq_train_labels, aes(x = total_cases)) +
  theme_bw() +
  ggtitle('Cases of dengue in Iquitos') +
  geom_histogram() +
  theme(plot.title = element_text(hjust = 0.5))
```


# Add total_cases column to *_train_features dataframes

```{r}
# sj_train_features <- left_join(sj_train_features, sj_train_labels)
sj_train_features$total_cases <- sj_train_labels$total_cases

# iq_train_features <- left_join(iq_train_features, iq_train_labels, by = c('city', 'year', 'weekofyear'))
iq_train_features$total_cases <- iq_train_labels$total_cases
```


# Correlation matrix

```{r}
m_sj_train_features <- data.matrix(sj_train_features)
m_sj_train_features <- cor(x = m_sj_train_features[,3:24], use = 'complete.obs', method = 'pearson')

m_iq_train_features <- data.matrix(iq_train_features)
m_iq_train_features <- cor(x = m_iq_train_features[,3:24], use = 'everything', method = 'pearson')
```

# Correlation Heatmap

```{r}
corrplot(m_sj_train_features, type = 'full', tl.col = 'black', method="shade")

corrplot(m_iq_train_features, type = 'full', tl.col = 'black', method = 'shade')
```


# Correlation Bar plot
```{r}

df_sj_train_features <- data.frame(m_sj_train_features)[2:21,] 
df_sj_train_features <- dplyr::select(df_sj_train_features, total_cases) 
                                    
df_iq_train_features <- data.frame(m_iq_train_features)[2:21,]
df_iq_train_features <- dplyr::select(df_iq_train_features, total_cases) 
```


```{r}
ggplot(df_sj_train_features, aes(x= reorder(rownames(df_sj_train_features), -total_cases), y = total_cases)) +
  geom_bar(stat = 'identity') +
  theme_bw() +
  ggtitle('Correlation of variables in San Juan') +
  ylab('Correlation') +
  xlab('Variables') +
  coord_flip()
```


```{r}
ggplot(df_iq_train_features, aes(x= reorder(rownames(df_sj_train_features), -total_cases), y = total_cases)) +
  geom_bar(stat = 'identity') +
  theme_bw() +
  ggtitle('Correlation of variables in Iquitos') +
  ylab('Correlation') +
  xlab('Variables') +
  coord_flip()

```


#Cleaning function step-by-step. 

# Function data cleaning

```{r}
data_clean <- function(df_dengue_features, df_dengue_labels = NULL, add_cases = TRUE) {
  
  # Filter by city
  sj_df_dengue_features <- filter(df_dengue_features, city == 'sj')
  iq_df_dengue_features <- filter(df_dengue_features, city == 'iq')
  
  if (add_cases == TRUE) {
  sj_df_dengue_labels <- filter(df_dengue_labels, city == 'sj')
  iq_df_dengue_labels <- filter(df_dengue_labels, city == 'iq')
  }
  # Removing week_start_date column
  sj_df_dengue_features <- dplyr::select(sj_df_dengue_features, -week_start_date)
  iq_df_dengue_features <- dplyr::select(iq_df_dengue_features, -week_start_date)

  # Fill of NA values with the previous value
  sj_df_dengue_features <- sj_df_dengue_features %>%
    do(na.locf(.))
  
  iq_df_dengue_features <- iq_df_dengue_features %>%
    do(na.locf(.))
  
  # Add total_cases to dataframe with features
  if (add_cases == TRUE) {
  sj_df_dengue_features$total_cases <- sj_df_dengue_labels$total_cases
  iq_df_dengue_features$total_cases <- iq_df_dengue_labels$total_cases
  }
  
  # Converting character columns into numbers
  sj_df_dengue_features <- as.data.frame(apply(sj_df_dengue_features,2,as.numeric))
  sj_df_dengue_features$city <- rep('sj', nrow(sj_df_dengue_features))
  iq_df_dengue_features <- as.data.frame(apply(iq_df_dengue_features,2,as.numeric))
  iq_df_dengue_features$city <- rep('iq', nrow(iq_df_dengue_features))
  
  result <- list(sj_df_dengue_features, iq_df_dengue_features )
  
  return(result)
}

```

# Development of model, prediction and comparison of predicted vs real total cases detected.

# Getting data_training clean

```{r}
data_train <- data_clean(train_features, train_labels, TRUE)
```


# Getting negative binomials models by city

```{r}
training_sj <- glm.nb(formula = total_cases ~ reanalysis_specific_humidity_g_per_kg +
                     reanalysis_dew_point_temp_k +
                     station_min_temp_c +
                     station_avg_temp_c, data = data_train[[1]])

training_iq <- glm.nb(formula = total_cases ~ reanalysis_specific_humidity_g_per_kg +
                        reanalysis_dew_point_temp_k +
                        station_min_temp_c +
                        station_avg_temp_c, data = data_train[[2]])
```


# Getting data_test clean

```{r}
data_test <- data_clean(test_features, add_cases = FALSE)
```



# Testing model with training data

```{r}
prediction_train_sj <-  predict(training_sj, data_train[[1]], type = 'response')
prediction_train_iq <-  predict(training_iq, data_train[[2]], type = 'response')

```

```{r}
df_prediction_train_sj <- data.frame('prediction' = prediction_train_sj, 'actual' = data_train[[1]]$total_cases,
                                     'time' = as.Date(train_features$week_start_date[1:936]))
df_prediction_train_sj <- melt(df_prediction_train_sj, id.vars = 'time')
ggplot(df_prediction_train_sj, aes(x = time, y = value, color = variable)) +
  geom_line() +
  ggtitle('Dengue predicted Cases vs. Actual Cases (City-San Juan) ')
```


```{r}
df_prediction_train_iq <- data.frame('prediction' = prediction_train_iq, 'actual' = data_train[[2]]$total_cases,
                                     'time' = as.Date(train_features$week_start_date[937:1456]))
df_prediction_train_iq <- melt(df_prediction_train_iq, id.vars = 'time')
ggplot(df_prediction_train_iq, aes(x = time, y = value, color = variable)) +
  geom_line() +
  ggtitle('Dengue predicted Cases vs. Actual Cases (City-Iquitos) ')
```


# Prediction of total_cases in the data set

```{r}
prediction_sj <-  predict(training_sj, data_test[[1]], type = 'response')
prediction_iq <-  predict(training_iq, data_test[[2]], type = 'response')
```

 
```{r}
data_prediction_sj <- data.frame('city' = rep('sj', length(prediction_sj) ), 
                                 'total_cases' = prediction_sj, 
                                 'weekofyear' = data_test[[1]]$weekofyear,
                                 'year' = data_test[[1]]$year )

data_prediction_iq <- data.frame('city' = rep('iq', length(prediction_iq) ), 
                                 'total_cases' = prediction_iq,
                                 'weekofyear' = data_test[[2]]$weekofyear,
                                 'year' = data_test[[2]]$year)

```

## Using the model for prediction

## Generate the submission

```{r}
submission_format$total_cases <- as.numeric(c(data_prediction_sj$total_cases, 
                                                   data_prediction_iq$total_cases))

submission_format$total_cases <- round(submission_format$total_cases, 0)
  
write.csv(submission_format,
          file = 'Model_2_Forecasted_Values.csv', row.names = F)

```


## Resources

 http://r-statistics.co/Poisson-and-Negative-Binomial-Regression-With-R.html
???	https://a-little-book-of-r-for-time-series.readthedocs.io/en/latest/
???	https://www.r-bloggers.com/search/time%20series/
???	https://rpubs.com/kaz_yos/pscl-2
???	https://www.analyticsvidhya.com/blog/2016/02/time-series-forecasting-codes-python/
???	https://www.digitalocean.com/community/tutorials/a-guide-to-time-series-visualization-with-python-3
???	https://www.digitalocean.com/community/tutorials/a-guide-to-time-series-forecasting-with-arima-in-python-3
???	http://www.blackarbs.com/blog/time-series-analysis-in-python-linear-models-to-garch/11/1/2016
???	

